```{r}
#load libraries
suppressMessages(library(rpart)) 
suppressMessages(library(rattle))
suppressMessages(library(rpart.plot))
suppressMessages(library(caret))
suppressMessages(library(forcats))
suppressMessages(library(dplyr))
suppressMessages(library(randomForest))
suppressMessages(library(lattice))
suppressMessages(library(naniar))
```
#EDA

Now that the data has been processed before creating regression models the relationship between variables and global sales will be explored to see if any features need transformations or if they should not be considered in the model.

```{r, fig.height = 3, fig.width = 5, fig.align = "center"}
suppressMessages(library(ggplot2))
suppressMessages(library(knitr))


#Create a boxplot of the response variable, sale price, to identify outliers
ggplot(vgData, aes(x=Global_Sales)) + geom_boxplot(col='blue') + labs(title= "Figure 2.11:", x= "Global Sales")
```
First, Figure 2.11 shows the box plot for global sales after the missing values were removed. There are now 13,838 different video-games and 36 columns.  There is one outlier which is the total sales for Wii sports. This outlier will not be removed since Wii sports is the most sold Nintendo game with over 82 million copies sold as it came with all Wii consoles. So it is not incorrectly input into the data set. There are still many games with global sales close to 0.1. This is also reasonable for some games that sold less than 10,000 copies. Again since the data was input by users and some of the sales for different platforms were missing the data a lot of the global sales listed are on the low side. For example the video game Minecraft is the most sold video game, however since it's PC sales were not included in the data set, which is its largest platform, the sales for Minecraft are incorrect and they are the tenth most sold on the list. Even with the incorrect data, we will see if we can use the information provided to predict the global sales available. With the data provided the mean of the Global Sales is 0.74 and the standard deviation is 2.20. Predictions will probably end up being on the low side because of this. The only other numeric feature is now year. 



```{r, fig.height = 4, fig.width = 6, fig.align = "center"}
#Plot year versus sales
ggplot(vgData, aes(x=Year, y=Global_Sales)) + geom_point() + labs(title= "Figure 2.12:", x= "Year", y ="Global Sales")
```

Figure 2.12 shows the relationship between Global Sales and Year. The relationship seems to be more quadratic then linear, so for any linear models a transformation will be used by adding a column of $Year^2$ to the features being considered. As years increase it seems so the does the distribution of global sales. However it may be important to consider the fact that games that been released earlier have been on the market for longer.

Next relationship between the categorical columns and Global Sales will be explored. 
 
```{r, fig.height = 20, fig.width = 30, fig.align = "center"}
#Plot a global sales versus a few categorical features
g1=ggplot(vgData, aes(x=Critic_Score,y=Global_Sales)) + geom_boxplot(fill = "white",col = 'black')+labs(title = 'Figure 2.13:', x ='Critic Score', y ='Global Sales')+ theme(text = element_text(size = 30))
g2=ggplot(vgData, aes(x=Developer,y=Global_Sales)) + geom_boxplot(fill = "white",col = 'black')+labs(title = 'Figure 2.14:', x ='Developer', y ='Global Sales')+ theme(text = element_text(size = 30))
g3=ggplot(vgData, aes(x=ESRB_Rating, y = Global_Sales)) + geom_boxplot(fill = "white",col = 'black') + labs(title = 'Figure 2.15:',x = 'ESRB_Rating', y = "Global Sales")+ theme(text = element_text(size = 30))
g4=ggplot(vgData, aes(x=Publisher,y=Global_Sales)) + geom_boxplot(fill = "white",col = 'black')+labs(title = 'Figure 2.16:', x ='Publisher', y ='Global Sales')+ theme(text = element_text(size = 30))
g5=ggplot(vgData, aes(x=Platform_Wii, y = Global_Sales)) + geom_boxplot(fill = "white",col = 'black') + labs(title = 'Figure 2.17:',x = 'Platform Wii')+ theme(text = element_text(size = 30))
g6=ggplot(vgData, aes(x=Genre, y = Global_Sales)) + geom_boxplot(fill = "white",col = 'black') + labs(title = 'Figure 2.18:',x = 'Genre')+ theme(text = element_text(size = 30))

gridExtra::grid.arrange(g1,g2,ncol=1,nrow=2)
```
Figure 2.13 shows the box plot for Critic Score against Global Sales. The Global Sales that seem to be the highest are the review scores that were high.  Figure 2.14 shows that Developer 'Other' had the most sales, this is consistent with the fact that most of the developers were put into the category other. 

```{r, fig.height = 20, fig.width = 40, fig.align = "center"}
#show plots for global sales versus categorical features
gridExtra::grid.arrange(g3,g4, ncol=1,nrow=2)
```

Figure 2.15 shows that the ratings E for Everyone and M for Mature have the highest global sales and EC for early childhood are not as popular. Figure 2.16 shows publishers with the most global sales seem to be Nintendo and Other while Microsoft and Bandai do not have a lot of sales.

```{r, fig.height = 20, fig.width = 30, fig.align = "center"}
#show plots for global sales versus categorical features
gridExtra::grid.arrange(g5,g6, ncol=1,nrow=2)
```
Lastly Figure 2.17 and Figure 2.18 show the relationship for the Platform Wii and Genre against Global Sales.  Most of the platforms generally show the same relationship with global sales, but Wii's have a larger distribution of global sales. The console Wii was one of the most popular Nintendo consoles before the Switch, so this is reasonable. For the category Genre, the genre's with the lowest sales where Educational, Sandbox, and Board game video games. The highest sales are in the categories Action, Misc, and Platform. 
